---
title: "Reproducible Research: Peer Assessment 1"
author: "Alexey Tkachev"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
We will use **dplyr** and **ggplot2** packages.  
```{r library_loading, echo=FALSE, results='hide'}
library(dplyr)
library(ggplot2)
```

```{r preprocessing}
df <- read.csv("activity.csv")
df$date <- as.Date(df$date)
df <- tbl_df(df)
```


## What is mean total number of steps taken per day?
At the first step we calculate the total number of steps taken per day and make a histogram of the total number of steps taken each day.
```{r steps_per_day}
tmp <- summarize(group_by(df, date), s=sum(steps, na.rm = TRUE))
days <- length(tmp$date)
ggp <- ggplot(tmp, aes(x = date, y = s/1000))
ggp <- ggp + geom_histogram(fill="gray", color=NA, stat = "identity")
ggp <- ggp + theme_bw()
ggp <- ggp + labs(title="Total number of steps taken per day", x="Date", y="Steps (thousands)")

```
 
```{r steps_per_day_draw, echo=FALSE, fig.width=9}
ggp
```

Now we calculate the mean and median of the total number of steps taken per day.
```{r mean_median}
steps.mean <- mean(tmp$s)
steps.median <- median(tmp$s)
```
Mean - `r steps.mean`  
Median - `r steps.median`  


## What is the average daily activity pattern?
```{r activity_pattern}
tmp <- summarize(group_by(df, interval), avg=mean(steps, na.rm = TRUE))
ggp <- ggplot(tmp, aes(x = interval, y = avg))
ggp <- ggp + geom_line(color="blue")
ggp <- ggp + theme_bw()
ggp <- ggp + labs(title="The average daily activity pattern", x="Intervals", y="Steps (average value)")
```

```{r activity_pattern_draw, echo=FALSE, fig.width=9}
ggp
```


## Imputing missing values
Now we will count the total number of missing values in the original dataset.
```{r imputing_values_count}
missing.rows <- is.na(df$steps)
missing.count <- sum(missing.rows)
```
Total number of missing values - `r missing.count`  

For the missing values we will use the median for that 5-minute interval.  
```{r imputing_values}
missing.rows <- which(missing.rows)
tmp <- summarize(group_by(df, interval), med=median(steps, na.rm = TRUE))
df.imp <- cbind(df, rep(tmp$med, days))
names(df.imp) <- c(names(df), "med")
df.imp$steps[missing.rows] <- df.imp$med[missing.rows]
df.imp <- select(df.imp, -med)

tmp <- summarize(group_by(df.imp, date), s=sum(steps, na.rm = TRUE))
ggp <- ggplot(tmp, aes(x = date, y = s/1000))
ggp <- ggp + geom_histogram(fill="gray", color=NA, stat = "identity")
ggp <- ggp + theme_bw()
ggp <- ggp + labs(title="Total number of steps taken per day (with imputed values)", x="Date", y="Steps (thousands)")
```

```{r imputing_values_draw, echo=FALSE, fig.width=9}
ggp
```

Now we calculate the mean and median values with imputed data
```{r mean_median_imp}
steps.mean.imp <- mean(tmp$s)
steps.median.imp <- median(tmp$s)
```
Mean with imputed data - `r steps.mean.imp`  
Median with imputed data - `r steps.median.imp` 

## Are there differences in activity patterns between weekdays and weekends?
