---
title: "Reproducible Research: Peer Assessment 1"
author: "Alexey Tkachev"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
We will use **dplyr** and **ggplot2** packages  
```{r, echo=FALSE, results='hide'}
library(dplyr)
library(ggplot2)
```

```{r}
df <- read.csv("activity.csv")
df$date <- as.Date(df$date)
```


## What is mean total number of steps taken per day?
At the first step we calculate the total number of steps taken per day and make a histogram of the total number of steps taken each day
```{r}
tmp <- summarize(group_by(df, date), s=sum(steps, na.rm = TRUE))
ggp <- ggplot(tmp, aes(x = date, y = s/1000))
ggp <- ggp + geom_histogram(fill="gray", color=NA, stat = "identity")
ggp <- ggp + theme_bw()
ggp <- ggp + labs(title="Total number of steps taken per day", x="Date", y="Steps (thousands)")

```
 
```{r, echo=FALSE, fig.width=9}
ggp
```

Now we calculate the mean and median of the total number of steps taken per day
```{r}
steps.mean <- mean(tmp$s)
steps.median <- median(tmp$s)
```
Mean = `r steps.mean`  
Median = `r steps.median`  


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
